---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Class project Overview
This project demonstrates the ability to create a document with embedded code.
For this example all code will be displayed but it is not necessary.
This project uses a data set provided on the Cousera web site, it was downloaded 
on May-14-2016 at this URL: 
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip


## Loading and preprocessing the data

* Read data into envioroment with "read.csv"
        (assumes data set is located in the current working directory)
* Create data set named "sum_Data" representing the total steps recorded each day using "tapply"
```{r}
myData <- read.csv("activity.csv", header = TRUE, sep = "," ,dec = "." )
sum_Data <- tapply(myData$steps, myData$date, sum)

```


## What is mean total number of steps taken per day?
Inspect the data set by making a Histogram to show the distribution of the total steps each day 

```{r, echo=TRUE}
hist(sum_Data, col = "blue", main = "Histogram of Total Steps per Day", xlab = "Number of Steps")
```

Calculate Mean and Median of the daliy toal steps
```{r, echo=TRUE}
meanData <- as.integer(mean(sum_Data, na.rm = TRUE))
medianData <- as.integer(median(sum_Data, na.rm = TRUE))

```
* The mean of the total number of steps taken per day is: `r meanData`
* The median of the total number of steps taken per day is: `r medianData`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
