---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Class project Overview
This project demonstrates the ability to create a document with embedded code.
For this example all code will be displayed but it is not necessary.
This project uses a data set provided on the Cousera web site, it was downloaded 
on May-14-2016 at this URL: 
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
The following libraies are used
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data

* Create data set named "myData" by reading data file into envioroment with "read.csv"
        (assumes data set is located in the current working directory)
    ```{r,echo = TRUE}
myData <- read.csv("activity.csv", header = TRUE, sep = "," ,dec = "." )
    ```
* Create data set named "sum_Data" representing the total steps recorded each day using "tapply"
    ```{r,echo = TRUE}
sum_Data <- tapply(myData$steps, myData$date, sum)
```

* Create data set named "interval_Data" of the average number of steps of each interval, the original data set "myData" is grouped by interval using "grouped_by", the varible names are given better descriptive names using "rename" , then the data set is created using summarize.

    ```{r,echo = TRUE}
interval_Data <- group_by(myData, interval) %>%
        rename(Average_Steps = steps, Interval = interval) %>%
        summarize( Average_Steps = mean(Average_Steps, na.rm = TRUE))
```


## What is mean total number of steps taken per day?
* Inspect the data set by making a Histogram to show the distribution of the total steps each day 

    ```{r,echo = TRUE}
hist(sum_Data, col = "blue", main = "Histogram of Total Steps per Day", xlab = "Number of Steps")
```

* Calculate Mean and Median of the daliy toal steps
    ```{r,echo = TRUE}
meanData <- as.integer(mean(sum_Data, na.rm = TRUE))
medianData <- as.integer(median(sum_Data, na.rm = TRUE))

```
* The mean of the total number of steps taken per day is: **`r meanData`**
* The median of the total number of steps taken per day is: **`r medianData`**

## What is the average daily activity pattern?
* Create a plot of the average number of steps taken during each interval
    ```{r,echo = TRUE}
g <- ggplot( data = interval_Data, aes( Interval, Average_Steps) )
g + geom_line(colour = "blue") + labs(title = "Time Series of Average Steps Taken") +
        labs(x = "Interval  [5 minutes]", y = "Number of Steps Taken")

```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
